MLExplain

### 2018-05-24 paper.tex
tiny fixes
Continue la sous-section sur les limitations

At first glance, JSExplain seems tightly linked to JavaScript. However, the
architecture of the project actually allows us to take off the interpreter
completely and replace it with another one.

@@

@@

% Toute une partie sur MLExplain lui-même...

>>>>

At first glance, JSExplain seems tightly linked to JavaScript. However, the
architecture of the project actually lets us completely replace the interpreter
with one for a different language.

@@

% Difference entre les constructeurs, raison principale de la traduction à la
% main

@@

Error management could not be done using exceptions. This is why we used monadic
computation. A monad is a data structure representing an action within a
context. A monad typically take the form of a type constructor. It provides the
following functions:

\begin{verbatim}
Let m, a monad:

val return : 'a -> 'a m
val bind : 'a m -> ('a -> 'b m) -> 'b m
\end{verbatim}

The function \texttt{return} let us build a new monadic value \verb|'a m| from
a value of type \verb|'a|. The function \texttt{bind} (or operator \texttt{>>=}) is
used to compose monads using a function of type \verb|'a -> 'b m|.

This signature is typical from functions that may fail -- for example
\verb|'a -> 'b option|. In this case, the \texttt{bind} function would apply
the supplied function on the monad only if represents a valid state (see figure
\ref{fig:error} for an implementation on \texttt{option}).

\begin{figure}[b]
  \begin{verbatim}
let bind_option opt f = match opt with
| Some value -> f value
| None -> None
  \end{verbatim}

  \label{fig:error}
  \caption{Implementation of operator bind on type option}
\end{figure}

Monads are an elegant way to manage errors, however it forces to nest many
levels of lambda expressions to chain multiple calls to functions that may
fail. This causes the source code to be extremely hard to read and therefore to
maintain. Our compiler supports \texttt{ppx} extensions, hence we implemented a
very simple syntax extension to transform a special let binding into a bind
function call:

\begin{verbatim}
(* This beautiful syntax *)
let%some a = function_that_could_fail 0 in
continuation a

(* becomes this *)
bind_option
  (function_that_could_fail 0)
  (fun a -> continuation a)
\end{verbatim}
